Emergence of the Southeast Asian islands as a driver for Neogene cooling

### 2020-07-31 Manuscript/Supplementary_Information.tex
update manuscript with the latest

\begin{flushleft}
{\Large \textbf{Supplementary Information for ``Emergence of the Southeast Asian islands as a driver for Neogene cooling''}}

Yuem Park\textsuperscript{1},
Pierre Maffre\textsuperscript{1},
Yves Godd\'eris\textsuperscript{3},
Francis A. Macdonald\textsuperscript{2},
Eliel A. Anttila\textsuperscript{2},
Nicholas L. Swanson-Hysell\textsuperscript{1}

@@

These supplementary information materials provide details on the model framework used in this study. The code for the GEOCLIM model used in this study can be found at: \url{https://github.com/piermafrost/GEOCLIM-dynsoil-steady-state}. The code that generated the inputs and analyzed the output of the GEOCLIM model can be found at: \url{https://github.com/Swanson-Hysell-Group/GEOCLIM_Modern}.

@@

Within the equation that governs the physical erosion rate in GEOCLIM (Equation 8), the values of the exponents ($m=0.5$ and $n=1$) are supported by compilations \citep{Lague2013a}. However, there is uncertainty in the value of the proportionality constant ($k_e$). We therefore tune this value such that the total erosion flux in the model under present-day slope and runoff conditions (see below) matches the total erosion flux estimated in \citet{Milliman2013a} ($20\times10^{12}$~kg/yr). Assuming that the density of eroded materials is 2500~kg/m$^{3}$, our tuned $k_e$ value is 0.0029110~m$^{1-m}$/yr$^{1-m}$.

@@

\begin{figure}[h!]
    \centering
    \includegraphics[width=0.7\textwidth]{Figures/paleogeographic_reconstructions.pdf}
    \caption{Paleogeographically-reconstructed paleoshorelines for SEAI at 5, 10, and 15~Ma, using the paleoshoreline reconstruction of this study coupled to the paleogeographic model of \citet{Matthews2016a}.}
    \label{fig:paleogeographic_reconstructions}
\end{figure}

@@

\begin{figure}[h!]
    \centering
    \includegraphics[width=1\textwidth]{Figures/pCO2_proxies.pdf}
    \caption{Proxy-based \pCOtwo estimates for the past 20~m.y. For data that were compiled within \citet{Foster2017a}, error bars indicate standardized uncertainties. For the \citet{Ji2018a} data, error bars represent the 16-84$^{th}$ percentile of resampled estimates. The \citet{Bereiter2015a} data come from the EPICA Dome C ice core.}
    \label{fig:pCO2_proxies}
\end{figure}

Proxy-based \pCOtwo estimates for the past 20~m.y. are shown in Figure \ref{fig:pCO2_proxies} \citep{Bereiter2015a, Foster2017a, Ji2018a}. Even when stringent quality criteria and the latest understanding of each of the \pCOtwo proxies have been applied to available \pCOtwo records, both significant uncertainty in the estimated \pCOtwo for any given data point as well as disagreement between techniques remain. Nevertheless, our modeled \pCOtwo values for 15~Ma (Fig. \ref{fig:scenario_pCO2_all}) resemble the higher end of proxy-based \pCOtwo estimates for the early-mid-Miocene. Given this scatter in the proxy-based \pCOtwo estimates, we instead infer the Neogene cooling trend from the Miocene benthic foram oxygen isotope record shown in Figure 1 in the main text.

>>>>

\begin{flushleft}
{\Large \textbf{Supporting Information for ``Emergence of the Southeast Asian islands as a driver for Neogene cooling''}}

Yuem Park\textsuperscript{1},
Pierre Maffre\textsuperscript{1},
Yves Godd\'eris\textsuperscript{3},
Francis A. Macdonald\textsuperscript{2},
Eliel S. C. Anttila\textsuperscript{2},
Nicholas L. Swanson-Hysell\textsuperscript{1}

@@

These supplementary information materials provide details on the model framework used in this study. The code for the GEOCLIM model used in this study can be found at: \url{https://github.com/piermafrost/GEOCLIM-dynsoil-steady-state/releases/tag/v1.0}. The code that generated the inputs and analyzed the output of the GEOCLIM model can be found at: \url{https://github.com/Swanson-Hysell-Group/GEOCLIM_Modern}.

@@

Within the equation that governs the physical erosion rate in GEOCLIM (Equation 8), the values of the exponents ($m=0.5$ and $n=1$) are supported by compilations \citep{Lague2013a}. However, there is uncertainty in the value of the proportionality constant ($k_e$). Using the simplification of a uniform value, we therefore tune this value such that the total erosion flux in the model under present-day slope and runoff conditions (see below) matches the total erosion flux estimated in \citet{Milliman2013a} ($20\times10^{12}$~kg/yr). Assuming that the density of eroded materials is 2500~kg/m$^{3}$, our tuned $k_e$ value is 0.0029110~m$^{1-m}$/yr$^{1-m}$.

@@

\begin{figure}[h!]
    \centering
    \includegraphics[width=0.6\textwidth]{Figures/paleogeographic_reconstructions.pdf}
    \caption{Paleogeographically-reconstructed paleoshorelines for SEAI at 5, 10, and 15~Ma, using the paleoshoreline reconstruction of this study coupled to the paleogeographic model of \citet{Matthews2016a}.}
    \label{fig:paleogeographic_reconstructions}
\end{figure}

@@

\begin{figure}[h!]
    \centering
    \includegraphics[width=1\textwidth]{Figures/pCO2_proxies.pdf}
    \caption{Proxy-based \pCOtwo estimates for the past 20~m.y. For data that were compiled within \citet{Foster2017a}, error bars indicate standardized uncertainties. For the \citet{Ji2018a} data, error bars represent the 16-84$^{th}$ percentile of resampled estimates. Individual data points from \citet{Cui2020a} were not published with uncertainties. The \citet{Bereiter2015a} data come from the EPICA Dome C ice core.}
    \label{fig:pCO2_proxies}
\end{figure}

Proxy-based \pCOtwo estimates for the past 20~m.y. are shown in Figure \ref{fig:pCO2_proxies} \citep{Bereiter2015a, Foster2017a, Ji2018a, Cui2020a}. Even when stringent quality criteria and the latest understanding of each of the \pCOtwo proxies have been applied to available \pCOtwo records, both significant uncertainty in the estimated \pCOtwo for any given data point as well as disagreement between techniques remain. Statistical methods that are often utilized in an attempt to extract trends from this data (e.g. locally-weighted scatter plot smoothing, LOWESS) evaluate the mean (with some weighting) of the highly scattered data in a given time interval. However, such methods are only appropriate for estimating the ``true'' value at any given time interval in datasets in which individual data points are being drawn from a single distribution about a mean value. The \pCOtwo proxy compilation shown in Figure \ref{fig:pCO2_proxies} consists of several distinct techniques and samples, each with their own set of assumptions and associated probability distributions that influence their reported values. Therefore, simply taking the mean of these values neglects the fact that some of these techniques/samples are more reliable others. Constraining how robust each \pCOtwo proxy technique is remains an important challenge, and as such the ``true'' \pCOtwo could plausibly lie anywhere within the full range of proxy-based \pCOtwo estimates for any given time interval. Nevertheless, our modeled \pCOtwo values for 15~Ma (Fig. \ref{fig:scenario_pCO2_all}) resemble the higher end of proxy-based \pCOtwo estimates for the early-mid-Miocene. Given this scatter in the proxy-based \pCOtwo estimates, we instead infer the Neogene cooling trend from the Miocene benthic foram oxygen isotope record shown in Figure 1 in the main text.
